### 2017-05-31 doc/intro.tex
intro updated

  We present novel methods for analyzing the activation patterns of
  RNNs from a linguistic point of view and explore the types of 
  linguistic structure they learn. As a case study, we use a multi-task gated
  recurrent network architecture consisting of two parallel pathways with
  shared word embeddings trained on predicting the
  representations of the visual scene corresponding to an input
  sentence, and predicting the next word in the same sentence. 
  We propose a method for estimating the amount of contribution
  of individual tokens in the input to the final prediction of the networks.   
  We show that the image prediction pathway: a) is sensitive to
  the information structure of the sentence, b) pays
  selective attention to lexical categories and grammatical functions
  that carry semantic information, and c) learns to treat the same
  input token differently depending on its grammatical functions in
  the sentence. In contrast the language model is comparatively more
  sensitive to words with a syntactic function. 
  Furthermore, we propose methods to explore the function
  of individual hidden units in RNNs and show that the two pathways
  of the architecture in our case study contain
  specialized units tuned to patterns informative
  for the task, some of which can carry activations to later time
  steps to encode long-term dependencies.

@@

\todo[inline]{once we finalise our experiments, we should update the following paragraphs and get rid of the macro/micro distinction.}
We report a thorough \emph{macro level} quantitative analysis to
provide a linguistic interpretation of the networks' activation
patterns. These experiments rely on a novel method we
call \emph{omission score} to measure the importance of input tokens 
to the final prediction of models that compute distributed representations 
of sentences. This can be applied to various RNN architectures, Recursive Neural Networks 
and Convolutional Neural Networks.\label{edit:introomissiongeneral}

Based on these omission scores our experiments show that the image
prediction pathway in general pays special attention to syntactic
categories which carry semantic content, and especially nouns. More
surprisingly, this pathway also learns learns to treat word types
differently depending on their grammatical function and their position
in the sequential structure of the sentence. In contrast, the textual
pathway and the standalone language model are especially sensitive to
the local syntactic characteristics of the input sentences.

@@

We also perform a \emph{micro level} analysis to explore \label{edit:introtopkgeneral}
the function of individual hidden units by introducing a method we dub \emph{top-k-contexts}.
It involves identifying the sentential contexts which yield the 
highest activation values and the analysis we present is applicable 
to uni-directional RNN architectures.
Through a qualitative examination of these contexts we
observe that in case of {\sc Imaginet} some of the contexts represent a particular 
syntactic category or dependency function, while others correspond to a semantic theme. 
We identify cases where the hidden units encode characteristics of 
the context that go beyond their lexical properties, and represent abstract 
patterns that are useful for the network's task. 
Furthermore, we explore and visualize units that carry their
activation over to later time steps to 
extract longer dependencies and more complex linguistic features. 
We also quantitatively show that features encoded by the language model are more associated 
with syntactic constructions than in case of the image prediction pathway.

>>>>

\todo[inline]{revise abstract}
  We present novel methods for analyzing the activation patterns of
  RNNs from a linguistic point of view and explore the types of 
  linguistic structure they learn. As a case study, we use a multi-task gated
  recurrent network architecture consisting of two parallel pathways with
  shared word embeddings trained on predicting the
  representations of the visual scene corresponding to an input
  sentence, and predicting the next word in the same sentence. 
  We propose a method for estimating the amount of contribution
  of individual tokens in the input to the final prediction of the networks.   
  We show that the image prediction pathway: a) is sensitive to
  the information structure of the sentence, b) pays
  selective attention to lexical categories and grammatical functions
  that carry semantic information, and c) learns to treat the same
  input token differently depending on its grammatical functions in
  the sentence. In contrast the language model is comparatively more
  sensitive to words with a syntactic function. 
  Furthermore, we propose methods to explore the function
  of individual hidden units in RNNs and show that the two pathways
  of the architecture in our case study contain
  specialized units tuned to patterns informative
  for the task, some of which can carry activations to later time
  steps to encode long-term dependencies.

@@

We report a thorough quantitative analysis to provide a linguistic 
interpretation of the networks' activation patterns. We present a 
series of experiments using a novel method we call \emph{omission score} 
to measure the importance of input tokens to the final prediction of models 
that compute distributed representations of sentences. 
Furthermore, we introduce a more global measure for estimating the 
informativeness of various types of n-gram contexts for each model.
These techniques can be applied to various RNN architectures, Recursive 
Neural Networks and Convolutional Neural Networks.

Our experiments show that the image prediction pathway in general 
pays special attention to syntactic categories which carry semantic 
content, and particularly nouns. More surprisingly, this pathway also 
learns to treat word types differently depending on their 
grammatical function and their position in the sequential structure of 
the sentence. In contrast, the textual pathway and the standalone 
language model are especially sensitive to the local syntactic 
characteristics of the input sentences.
\todo[inline]{say something about the findings of the last experiment.}

@@

%We also perform a \emph{micro level} analysis to explore \label{edit:introtopkgeneral}
%the function of individual hidden units by introducing a method we dub \emph{top-k-contexts}.
%It involves identifying the sentential contexts which yield the 
%highest activation values and the analysis we present is applicable 
%to uni-directional RNN architectures.
%Through a qualitative examination of these contexts we
%observe that in case of {\sc Imaginet} some of the contexts represent a particular 
%syntactic category or dependency function, while others correspond to a semantic theme. 
%We identify cases where the hidden units encode characteristics of 
%the context that go beyond their lexical properties, and represent abstract 
%patterns that are useful for the network's task. 
%Furthermore, we explore and visualize units that carry their
%activation over to later time steps to 
%extract longer dependencies and more complex linguistic features. 
%We also quantitatively show that features encoded by the language model are more associated 
%with syntactic constructions than in case of the image prediction pathway.
